\documentclass{article}
\input{../homework.sty}

\title{Homework 4}
\author{Austin Gill}
\date{April 23, 2019}

\begin{document}
\maketitle

The following problems forbid using Example 11.3. Before we can proceed, we must make several definitions.

\begin{defn}
    A grammar $G = (V, T, S, P)$ is said to be \textbf{linear} if every production is either right-linear or left-linear. That is, every production has one of the following forms
    \begin{align*}
        A & \to xB \\
        A & \to Bx \\
        A & \to x
    \end{align*}
    For $A$, $B$, in $V$ and $x \in T^*$.
\end{defn}

\begin{defn}
    A grammar $G = (V, T, S, P)$ is \textbf{context-free} if all productions in $P$ have the form \[A \to x\] where $A \in V$ and $x \in {(V \cup T)}^*$.
\end{defn}

\begin{defn}
    A context-free language $L$ is said to be \textbf{linear} if there exists a linear context-free grammar $G$ such that $L = L(G)$.
\end{defn}

\begin{defn}
    A language $L$ is said to be \textbf{deterministic context-free} if and only if there exists a deterministic pushdown accepter $M$ such that $L = L(M)$.
\end{defn}

\section{}
\begin{quote}
    Find two examples of languages that are linear but not deterministic context free.
\end{quote}

$L = \{ww^R\}$ is nondeterministic, but linear.\todo{Show that it's linear.}

\todoinline{Find another language and show that it is linear.}

\section{}
\begin{quote}
    Find two examples of languages that are deterministic context free but not linear.
\end{quote}

The following three languages are not linear.
\begin{align}
    L & = \{ a^n b^n a^m b^m \mid n \geq 0, m \geq 0 \}\label{eq:L-cat-L} \\
    L & = \{ w \in {\{a, b\}}^* \mid n_a(w) \leq n_b(w) \}                \\
    L & = \{ w \in {\{a, b, c\}}^* \mid n_a(w) + n_b(w) = n_c(w) \}
\end{align}

Notice that \autoref{eq:L-cat-L} is the linear language $L = \{a^n b^n \mid n \geq 0 \}$ concatenated with itself.

\todoinline{Show that these languages are deterministic context free.}
\end{document}
